Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm

### 2019-09-27 paper/virtual_cytoplasm_v2.tex
saving tex changes

While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.  Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluoroscence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way. Another essential class of methods come from computational modeling and simulation. By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now ~\cite{vanGunsteren2006a}.

As the available computational power increases, models become increasingly complex and are better representations of reality. The most simple class of computational models apply brownian dynamics calculations to probe macromolecules imersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}. These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}.More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}. However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm. This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms ~\cite{feig2015complete}. This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations ~\cite{yu2016biomolecular}. Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.

@@

\subsection*{Initial structures}
The proteins and tRNA were downloaded from Protein Data Bank (PDB). We looked for the proteins structures that were either from or expressed in {\em E-coli}. In case of 1U22 (MetE) and 2EIP (Ppa) we used a loop-closure modelng tool based on Random Coordinate Descent (RCD) method~\cite{Chys2013} to correct the information for missing residues. The four metabolites were parametrized using GAFF and Antechamber.

@@

\subsubsection{Protein fraction}
We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{Link1997}. The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in {\em E. coli}, respectively, while the other six proteins contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}). Their crystallogaphic structures were used to insert them in the cytoplasm model. The number of copies of each protein in the model was calculated as the rounded up fold increase in their abundance when compared to the least abundant protein, which was set to have the copy number of 1. Then, this number was divided by the oligomeric state of each protein. This was done because the original oligomeric state for each protein was kept as reported in their crystallogaphic structure.

@@

\subsubsection{Inorganic ions fraction}
Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and ATP. K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substution of randomly selected water molecules.

@@

In order to do this, the droplets around each component are also constructed iteractively. The number of water molecules that we must place in the simulation box is known from the desired biomolecular fraction, which is a parameter of the model, and the total biomolecular mass of the system. However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules. Since $l$ depends on a series of factors such as the shape, size and abundancy of each element, we define it iteratively. We start by taking a droplet of thickness $l$ = 3 \r{A} around all elements and counting the number of water molecules that we would add to the box with such droplets. If it is smaller than the number we need, $l$ is increased by $dl$ = 1 \r{A}. If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increate $l$ with the new smaller $dl$. We can carry this process until an arbitrary precision cutoff, such as 5\%, is satisfied.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking_volume.pdf} 
\includegraphics[scale=0.11]{box_big.png}
\includegraphics[scale=0.11]{box_small_.png}
%\includegraphics[scale=0.5]{shrinking_density.pdf}

@@

\caption{A) Interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). \st{The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica. The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.} The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica. The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundancy of the protein minus the sasa of the whole protein averaged over the number of replica and the abundancy of the protein. The error bars show the standard errors. B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.}
\label{fig:structural_integrity_chain}
\end{figure}

@@

\subsubsection{Structural dynamics}
Proteins are slighly more flexible in the crowded environment. Root mean square fluction (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ... as shown before (Figure~\ref{fig:rmsf}). This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected. In addition to that, our results show that the proteins are slighly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}). This is likely due to the nonspecific interations and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.

@@

The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders. However, tRNA is not causing such phenomenon. Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$. tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few "sources" of Mg$^{2+}$ in the whole system. We looked for ways of avoind aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).

We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure~\ref{fig:avoiding_aggregation}). While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}.

@@

@@

The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A). The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules. We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement. Thus, tRNA is considered as an outlier in our cytoplasm model. In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data. Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E. coli}, as measured by fluorescence recovery after photobleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.

>>>>

While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.  Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way. Another essential class of methods come from computational modeling and simulation. By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now ~\cite{vanGunsteren2006a}.

As the available computational power increases, models become increasingly complex and are better representations of reality. The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}. These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}. More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}. However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm. This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms ~\cite{feig2015complete}. This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations ~\cite{yu2016biomolecular}. Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.

@@

\subsection*{Initial structures}
The proteins and tRNA were downloaded from Protein Data Bank (PDB). We looked for the proteins structures that were either from or expressed in {\em E-coli}. In case of 1U22 (MetE) and 2EIP (Ppa) we used a loop-closure modeling tool based on Random Coordinate Descent (RCD) method~\cite{Chys2013} to correct the information for missing residues. The four metabolites were parametrized using GAFF and Antechamber.

@@

\subsubsection{Protein fraction}
We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{Link1997}. The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in {\em E. coli}, respectively, while the other six proteins contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}). Their crystallographic structures were used to insert them in the cytoplasm model. The number of copies of each protein in the model was calculated as the rounded up fold increase in their abundance when compared to the least abundant protein, which was set to have the copy number of 1. Then, this number was divided by the oligomeric state of each protein. This was done because the original oligomeric state for each protein was kept as reported in their crystallographic structure.

@@

\subsubsection{Inorganic ions fraction}
Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and ATP. K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substitution of randomly selected water molecules.

@@

In order to do this, the droplets around each component are also constructed iteratively. The number of water molecules that we must place in the simulation box is known from the desired biomolecular fraction, which is a parameter of the model, and the total biomolecular mass of the system. However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules. Since $l$ depends on a series of factors such as the shape, size and abundance of each element, we define it iteratively. We start by taking a droplet of thickness $l$ = 3 \r{A} around all elements and counting the number of water molecules that we would add to the box with such droplets. If it is smaller than the number we need, $l$ is increased by $dl$ = 1 \r{A}. If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increase $l$ with the new smaller $dl$. We can carry this process until an arbitrary precision cutoff, such as 5\%, is satisfied.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking_volume.pdf} 
\includegraphics[scale=0.5]{box.pdf} 
% \includegraphics[scale=0.11]{box_big.png}
% \includegraphics[scale=0.11]{box_small_.png}
%\includegraphics[scale=0.5]{shrinking_density.pdf}

@@

\caption{A) Chain interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica. The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundance of the protein minus the sasa of the whole protein averaged over the number of replica and the abundance of the protein. The error bars show the standard errors. B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.}
\label{fig:structural_integrity_chain}
\end{figure}

@@

\subsubsection{Structural dynamics}
Proteins are slightly more flexible in the crowded environment. Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ... as shown before (Figure~\ref{fig:rmsf}). This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected. In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}). This is likely due to the nonspecific interactions and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.

@@

The initial detection of aggregation in our cytoplasm model was due to the out-lier behavior of the translational diffusion constant of tRNA when compared to the other crowders. However, tRNA is not causing such phenomenon. Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$. tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few "sources" of Mg$^{2+}$ in the whole system. We looked for ways of avoid aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).

We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure~\ref{fig:avoiding_aggregation}). While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Figure~\ref{fig:protonation}).

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_mg_metabolites.pdf}
\caption{A) Radial distribution function (rdf) showing the probability of finding an ATP$^{-3}$ or an ATPH\textsubscript{3} around Mg$^{2+}$. B) Radial distribution function (rdf) showing the probability of finding an FBP$^{-4}$ or an FBPH\textsubscript{4} around Mg$^{2+}$. }
\label{fig:protonation}
\end{figure}

@@

The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A). The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules. We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement. Thus, tRNA is considered as an out-lier in our cytoplasm model. In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data. Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E. coli}, as measured by fluorescence recovery after photo-bleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.
